%!TEX root = main.tex

\section{The SIS-LWE Framework}

\begin{frame}
\frametitle{Construction of $q$-ary lattice (Primal / Construction A)}

Let $q$ be a prime\footnote{Not necessarly, but simpler.} integer, and $n < m$ two positive integers.

The matrix $\vec A \in \Z_q^{m \times n}$ spans the $q$-ary lattice:
\begin{align*}
\Lambda_q(\vec A) := & \,\{\vec x \in \Z^m \, | \, \exists \vec y \in \Z_q^n, \, \vec x \equiv \vec A \vec y \bmod q\}  \\
= & \,\, \vec A \cdot \Z_q^n + q \Z^m
\end{align*}
\begin{exampleblock}{Lattice parameters}
Assuming $\vec A$ is full-rank:
\begin{itemize}
  \item $\dim(\Lambda_q(\vec A)) = m$
  \item $\vol(\Lambda_q(\vec A)) = q^{m - n}$
\end{itemize}
\end{exampleblock}
\end{frame}

\begin{frame}
\frametitle{Construction of $q$-ary lattice (Dual / Parity-Check)}

Let $q$ be a prime\footnote{Not necessarly, but simpler.} integer, and $n < m$ two positive integers. \\
The matrix $\vec A^{\rred t} \in \Z_q^{\rred{n \times m}}$ is the parity-check of the lattice:
\begin{align*}
\Lambda_q^\bot(\vec A^{\rred t}) := & \,\{\vec x \in \Z^m \, | \vec A^{\rred t} \vec x \equiv \vec 0 \bmod q\}  \\
= & \ker (\vec x \mapsto \vec A^{\rred t} \vec x \bmod q) \phantom{\Z_q^n}
\end{align*}
\begin{exampleblock}{Lattice parameters}
Assuming $\vec A$ is full-rank:
\begin{itemize}
  \item $\dim(\vec A) = m$
  \item $\vol(\vec A) = q^{\rred n}$
\end{itemize}
\end{exampleblock}
\end{frame}


\begin{frame}
\frametitle{The Short Integer Solution Problem (SIS)}
\begin{definition}[SIS assumption]
  Given a random matrix $\vec A$ \\
  Finding a small non-zero $\vec x \in \Z_q^n$ such that $\vec A \vec x \equiv \vec 0 \bmod q$ is {\bf hard}. \\
\end{definition}
\pause
\begin{exampleblock}{Lattice formulation}
  Solving Approx-SVP in $\Lambda_q^\bot(\vec A^t)$ is {\bf hard}.
\end{exampleblock}

Worst-case to average case connection due to [Ajtai 1998].

\end{frame}

\begin{frame}
\frametitle{Simple application of SIS}
Set $\mathcal S = \{0, 1\}^m$ and consider the function:
\[
  f_{\vec A}: \mathcal S \rightarrow \Z_q^n, \qquad \vec x \mapsto \vec A^t \vec x \bmod q
\]

\begin{exampleblock}{SIS $\Rightarrow$ Collision Resistant Hashing and One-Way Function}
  \begin{itemize}
    \item Finding collision\footnote{Collision must exist whenever $m > n \log_2 q$} is as hard as SIS \hfill {\scriptsize (take the difference)}
  \end{itemize}
  Moreover, if $m \gg n \log q$:
  \begin{itemize}
    \item $f_{\vec A}$ is highly surjective \hfill {\scriptsize (many pre-images exists)}
    \item Finding pre-images is hard.
  \end{itemize}
\end{exampleblock}
\end{frame}

\begin{frame}
\frametitle{The Learning With Error problem (LWE)}
Let $\chi$ be a distribution of small errors $\ll q$. 
\begin{definition}[Decisional LWE]
For $\vec A \gets \Z_q^{m \times n}$, $\vec s \gets \Z_q^{n}$, $\bblue{\vec e} \gets \chi^m$, \\ distinguishing $(\vec A, \vec A \vec s + \bblue{\vec e})$ from uniform is hard.
\end{definition}
\begin{definition}[Search LWE]
For $\vec A \gets \Z_q^{m \times n}$, $\vec s \gets \Z_q^{n}$, $\bblue{\vec e} \gets \chi^m$, \\ given $(\vec A, \vec A \vec s + \bblue{\vec e})$, finding $\vec s$ is hard.
\end{definition}
Both problems are easily proved equivalent.
\pause
\begin{exampleblock}{Lattice formulation}
  Solving BDD in $\Lambda_q(\vec A)$ is {\bf hard}.
\end{exampleblock}
Worst-case to average case connection due to [Regev 2005].
\end{frame}

\begin{frame}
\frametitle{LWE as unique-SVP (The embedding technique)}
Given $(\vec A, \vec b = \vec A \vec s + \bblue{\vec e})$, consider

\[\Lambda = \Lambda_q(\vec A, \vec b) \]

Then:
\begin{itemize}
  \item $\bblue {\vec e} \in \Lambda$, and $\| \bblue {\vec e} \| \approx \sigma \sqrt m$
  \item one would expect $\lambda_1(\Lambda) \approx \sqrt{\frac m {2\pi e}} \cdot q^{1-n/m}$
\end{itemize}

\begin{exampleblock}{Alternative lattice formulation}
  Solving Unique-SVP in $\Lambda_q(\vec A, \vec b)$ is {\bf hard}.
\end{exampleblock}

\end{frame}


\begin{frame}
\frametitle{Simple application of LWE}
Set $\mathcal S = \{-\sigma, \dots \sigma\}^m$ and consider the function:
\[
  g_{\vec A}: \Z_q^n \times \mathcal S \rightarrow \Z_q^m, \qquad (\vec s, \bblue{\vec e}) \mapsto \vec A \vec s + \bblue{\vec e} \bmod q
\]

\begin{exampleblock}{LWE $\Rightarrow$ Secret-Key Encryption}
  Idea : Noisy one-time pad
  \begin{itemize}
    \item $Enc_{\vec s}(m\in\{0,1\} ) = (\vec a,\vec a^t \vec s + e + \lfloor \frac q 2 \rceil m)$ 
    \item $Dec_{\vec s}(\vec a, b ) = \lfloor \frac 2 q (b - \vec a^t \vec s) \rfloor \bmod 2$ 
  \end{itemize}
\end{exampleblock}
\end{frame}
