%!TEX root = main.tex

\section{Encryption is easy}

\begin{frame}
\frametitle{Encryption is easy}

{\bf Idea:}
\begin{itemize}
  \item Use one short lattice vector {\hfill \scriptsize(rather than a full good basis $\bblue{\vec B}$)}
  \item This short vector is easy to hide: LWE as unique-SVP
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Public Key Encryption, [Regev 2005]}
$m \gg n \log q$.
\begin{itemize}
  \item $SK = \vec s \in \Z_q^m$
  \item $PK = (\vec A; \vec b = \vec A \vec s + \bblue{\vec e}) \in \Z_q^{(n+1) \times m}$
  \item $Enc(m) = (\bblue{\vec t^t} \cdot \vec A, \bblue{\vec t^t} \cdot \vec b + \lfloor \frac q 2 \rceil m + e)$, where $\vec t \gets \{0,1\}^{n+1}$
  \item $Dec(\vec x^t, y)$ Compute \[d = y - \vec x^t \vec s = \bblue{\vec t^t} \bblue{\vec e} + e + \lfloor \frac q 2 \rceil m\] and return
   $m = \lfloor \frac 2 q d \rfloor \bmod 2$
\end{itemize}
\pause
\begin{exampleblock}{Proof sketch for CPA security}
\begin{itemize}
  \item Replace PK by uniform random $(\vec A, \vec b)$
  \item Apply the left-over hash lemma on $\vec t$ over $(\vec A, \vec b)$
  \item $Enc(m)$ is statistically close to uniform.
\end{itemize}
\end{exampleblock}
\end{frame}

\begin{frame}
\frametitle{PKE / Approx. Key-Exchange [Lindner Peikert 2011]}
Using a Systematic-Normal form, one can assume that $\bblue{\vec s} \gets \chi^n$ is small as well.
Take $m = n$.

\begin{itemize}
  \item $PK = \vec s \in \Z_q^n$
  \item $SK = (\vec A; \vec b = \vec A \bblue{\vec s} + \bblue{\vec e}) \in \Z_q^{(n+1) \times n}$
  \item $Enc(m) = (\vec A^t \bblue{\vec s'} + \bblue{\vec e'}, \vec b^t \bblue{\vec s'} + \bblue{\vec e'} + \bblue e + \lfloor \frac q 2 \rceil m)$

  \item $Dec(\vec x, y):$ Compute \[d = y - \vec x^t \vec s = \bblue{\vec s^t} \bblue{\vec e'} + \bblue{\vec s'^t} \bblue{\vec e} + e + \lfloor \frac q 2 \rceil m\] and return
   $m = \lfloor \frac 2 q d \rfloor \bmod 2$
\end{itemize}
\pause
\begin{exampleblock}{Proof sketch for CPA security}
\begin{itemize}
  \item Replace PK by uniform random by LWE assumptuion
  \item Replace $Enc(m)$  by uniform random by LWE assumptuion
\end{itemize}
\end{exampleblock}
Can also be made an approximate key Exchange.
\end{frame}


\begin{frame}
\frametitle{Chosen-Ciphertext Secure ?}

Are the above CCA-secure ?
\begin{alertblock}{NO !}
It is Additively Homomorphic therefore can't be CCA2. \\
CCA1 attacks left as an exercise.\footnote{Toy with the error and see if Dec. fails}
\end{alertblock}

Generic Transform to CCA security in the Random Oracle Model ?
\pause
\begin{exampleblock}{Yes [Peikert 2013]}
Correctness needs to hold with overwhelming probability. 
\end{exampleblock}
\pause
And in the plain Model ?
\begin{exampleblock}{Yes}
But costly: requires Trapdoors (e.g [Micciancio Peikert 2012])
Open question: Cramer-Shoup for lattices ?
\end{exampleblock}



\end{frame}